Environmental stress affects niche breadth in plant-pollinator communities

### 2019-11-17 paper/manuscript.Rmd
Regression to the mean idea
Check spelling manuscript

Species interactions are determined in part by niche processes (the matching of traits) and partly by neutral processes (more abundant species are more likely to encounter each other and, thus, interact). 
The environment can influence both of these processes. 
It is, therefore, not surprising that, despite limitations on the spatial extent or the number of environmental gradients considered, multiple studies have been able to show how changes to interactions can be related to environmental change [@tylianakis_ecological_2017].
For instance, some studies suggest that the strenghth of some trophic interactions, like predation [@mckinnon_lower_2010; @vucic-pestic_warming_2011] and herbivory [@baskett_latitudinal_2018], can increase with temperature but might decrease with precipitation [@pires_predicted_2016]. 
Some other studies, however, have shown either no effect (on average) or non-linear effects of temperature or precipitation on plant-pollinator interactions [@devoto_patterns_2005; @gravel_bringing_2018].
Overall, while it looks clear that pairwise interactions respond to environmental drivers, there is high variability in the response [@tylianakis_global_2008].

@@

Here, we investigate whether, and how, environmental stress can systematically affect species degree. 
Our main aim is to test the two competing hypotheses that relate environmental stress and species degree and investigate whether this changes across species or between trophic guilds. 
We propose that specialist species can become "facultative" generalists to reduce their vulnerability to the absence of preferred partners [for example, when variations in climate decouple phenologies; @benadi_specialization_2014]. 
We therefore also expect that as environmental stress increases species with a relatively small number of partners are more likely to engage with more partners and broaden their trophic niche. 
Species with a large number of partners, on the other hand, should have a larger pool of available partners and might, therefore, be more likely to narrow their niche under environmental stress by focusing on the most benefitial partners. 
Importantly, when testing these hypotheses, we control for the potential effects of the environment in community composition and the size of the species fundamental niche, both from an Eltonian (interactions) and Grinellian (environment) perspective.

@@

We retrieved plant-pollinator networks from the Web of Life database [@fortuna_web_2014]. 
This database contains datasets originating from `r length(interaction_citations$BIBTEXKEY)` studies published in the primary literature between `r int_citations_year_range[1]` and `r int_citations_year_range[2]`. 
Calculating the environmental stress of species in their community and their Eltonian niche breadth required us to reduce both the taxonomic and distributional/locational uncertainty.
A critical step towards reducing this uncertainty is to ensure that the names used to identify species are valid and unambiguous, which in turn allow us to obtain further information from biological databases and accurately match species across studies. 
Therefore, our first step was to ensure consistent spelling and standarisation of species names synonyms (See Supplementary Methods).
This ensured that the matching of species across studies was as accurate as possible. 
The cleaning process resulted on a total of `r format(stats_clean_int$pla_id, big.mark = ",")` plants and `r format(stats_clean_int$ani_id, big.mark = ",")` pollinator species distributed across `r format(stats_clean_int$loc_id)` locations arround the globe (Figure \@ref(fig:fig-worldmap), \@ref(fig:fig-distribution-species-mult-locations)).

@@

We retrieved `r round(sum(n_occurrences$N)/1000000, digits = 1)` million occurrences from the Global Biodiversity Information Facility (GBIF; https://www.gbif.org). 
Issues with data quality are a central issue hampering the use of publicly available species occurrence GBIF data in ecology and biogeography (Jetz et al 2019). 
We, therefore we followed a series of filters and geographic heuristics to correct or remove erroneous and imprecise referencing records [See supplementary methods; @zizka_coordinatecleaner_2019] which allowed us to identify and remove `r round(sum(n_dirty_occurrences$N)/1000000, digits = 1)` million problematic occurrences from further analysis.
We integrated the occurrences from our plant-pollinator communities to the cleaned occurrences retrieved from GBIF.

@@

Sets of occurrence data tend to be spatially aggregated due to sample bias (tendency to collect close to cities, certain countries). 
Spatial autocorrelation arises in ecological data because geographically clumped records tend to be more similar, in physical characteristics and/or species abundances, than are pairs locations that are farther apart. 
To account for such spatial dependency in occurrence data, if a species had more than one occurrence records within one of the cells of the bioclimatic raster, we only included one of the occurrence records. 
We did this to avoid giving more weight to areas with a high number of occurrences, a common scenario in occurrence records collected opportunistically as the ones we use here. 
In this step we removed `r scales::percent(percent_thinned)` of the occurrences which resulted in a total of `r round(sum(n_thinned_occurrences$N)/1000000, digits = 1)` million occurrences used in our niche analysis.

@@

We then used a set of bayesian multilevel models to evaluate the impact of environmental suitability on the number of partners a species has. 
Specifically, we use the normalised degree of species as our response variable, this is, the number of species it interacts with given the number of species in the opposite guild [@martin_gonzalez_centrality_2010]. 
The normalised degree was modelled using a logit link function, and a binomial distribution in which the number of species interacts with is the number of successes, and the number of species in the opposite guild is the number of trials. 
We are aware that whether species interact or not is not a Bernoulli process as species interactions are not strictly independent from each other. 
However, a binomial distribution allows us to account for the differences in species richness across communities indirectly. 
Importantly, however, results are qualitatively similar when we model species degree directly using a Poisson distribution and a logarithmic link function.

We evaluate four models to assess the relative importance of suitability. 
A first model, our baseline model, included three population-level predictors and two grouping levels, species and the community.
The population-level predictors in the baseline model, commonly called fixed effects, were the habitat suitability, the species guild (plant or a pollinator), and its number of known possible partners.
We included the number of known possible partners as a predictor in our models as it allows us to control for the environmental effects on species co-occurrence. 
We calculate this metric by determining the number of partners with which the species is known to interact in any other community.
Controlling for the number of potential partners makes our model a particularly stringent test of our environmental stress hypotheses because this variable could explain a large proportion of variance. 
Often the potential and the actual number of partners is the same or very close to each other, especially for rare species present only in a few communities.
As we were interested in understanding whether the effect of habitat suitability is conditional on the species guild (plant or pollinator), we, therefore, included guild and its interaction with suitability in the model. 
We allowed the intercept of degree and slope of the suitability-degree relationship to vary among species, this allowed us to investigate two questions. 
First, it allows us to inspect the extent to which suitability is a population or a group level effect. 
Second, by investigating the correlation between the intercept and the slope as a model parameter, it allowed us to inspect the extent by which species with a small or large number of interactions respond to increasing levels of environmental stress. 
To account for unmeasured differences between communities, like sampling effort, sampling method, or diversity, we also calculated an intercept for each community in our study.
To facilitate model interpretation and convergence, we scaled all continuous variables to have a mean of zero and a unit standard deviation.

@@

After performing our sensitivity analysis, we found that we need roughly `r readd(min_occurrences_factor)` independent occurrences for each community for which we calculated a suitability value in order to obtain a mean absolute error below `r config$min_suitability_error` (Fig. \@ref(fig:fig-sensitivity-analysis)). 
We therefore removed from further analyses `r n_spp_not_enough_occ` species for which we did not have enough occurrences to obtain robust estimates.
When inspecting the suitability values of the analysed species we found that most communities included species for which habitat suitability was low and species for which it was high (Fig. \@ref(fig:fig-median-suitability)).

@@

Our models performed relativelly well.
The bayesian R-squared for our baseline model was `r round(bayesian_r2_baseline[1], digits = 2)`, which indicates our models were able to capture a large proporton of the variability on the data. 
Overall we found that environmental suitability does not show a consistent pattern accross species. 
Indeed, when looking at the population level effects, suitability has virtually no relationship with normalised degree, neither for plants or pollinators (Figure \@ref(fig:fig-conditional-effects)b). 
However, suitability is an important predictor as the WAIC difference between our baseline model and that that did not include suitability was `r round(waic_diff_suitability)` ± `r round(waic_se_diff_suitability)` (Table \@ref(tab:tab-model-comparison)).
This apparent discrepancy can be explained by the variability of the suitability-degree relationship across species.

@@

fig_conditional_effects_legend <- 
  paste0("Conditional effects of predictors in our baseline model. The predicted values of the number of interacting species are based on a hypothetical community with ",
         median_n_pla, " plants and ", 
         median_n_ani, " pollinators",
         ". These values correspond to the median number of species in each guild respectively. In each panel we condition on the mean values of all other predictors in the model, mean values for each predictor are indicated with a vertical dashed line. For model fitting, we scaled all predictors to have mean of zero and unit variance. However, with the exception of environmental niche size, here we show the unscaled predictors to facilitate interpretation. To illustrate the uncertainity around the fitted estimates we plot the fits of 100 independent draws from the posterior distribution. The thick lines indicate the mean values of the response distribution. As there was no interaction between guild and generality or the number of possible interactions, for these two predictors we only show conditional the conditional effect of pollinators.")

@@

While for some species there is a strong negative relationship between suitability and normalised degree, for some others with a strong positive relationship (Figure \@ref(fig:fig-random-effects)a). 
Interestingly, the slope this relationship was negatively correlated with the species' intercept in the model (Figure \@ref(fig:fig-random-effects)b). 
The mean correlation coefficient was `r round(random_cor$Estimate, 2)` [`r round(random_cor$Q2.5, 2)`, `r round(random_cor$Q97.5, 2)`].
In other words, the slope of the suitability-degree relationship was more likelt to be positive for species with a smaller number of partners in lower-suitability and negative for species with a larger number of partners in lower-suitability conditions.

@@

fig_random_effects_legend <- 
  paste0("Species level effects of suitability. (a) Conditional effect of suitability for individual species. To facilitate visualization we show only species for which there is suitability information in at least six communities (", 
         n_random_pla, " plants and ", 
         n_random_ani, " pollinators). As in the previous figure, fitted values assume a hypothetical community of median size. In each panel we highlight two species for which the relationship between environental suitability and normalised degree was particularly strong. ", 
         "(b) The correlation between the species' intercept and the species' slope of suitability was negatively correlated.")

@@

As expected, we found a strong and positive relationship between the number of possible interactions and the number of realised interactions in the community. 
There was a large difference on WAIC between the model that included this predictor and that that excluded it. 
This result indicates that the availability of potential partners accounts for a large proportion of the variabiity on species degree.  
Importantly, our findings about the suitability-degree relationship were qualitatively unchanged whether this variable was included or not.

The group level variation among communities was larger than that among species which further indicates the importance of the local context when determining species degree.
Specifically, the standard deviation (in the parameters scale) of the community intercepts was `r round(sd_loc_id$Estimate, 2)` [`r round(sd_loc_id$Q2.5, 2)`, `r round(sd_loc_id$Q97.5, 2)`] while the standard deviation of the species intercept was `r round(sd_org_id$Estimate, 2)` [`r round(sd_org_id$Q2.5, 2)`, `r round(sd_org_id$Q97.5, 2)`], and that of the species' suitability slope was `r round(sd_org_id_suitability$Estimate, 2)` [`r round(sd_org_id_suitability$Q2.5, 2)`, `r round(sd_org_id_suitability$Q97.5, 2)`] (95% credible intervals shown within square brakets).

@@

We set out to explore wether and how environmental stress can systematically affect species degree. 
After accounting for the pool of potential partners, we found that indeed environmental conditions contribute to determining wether a species is a generalist or a specialist in their community. 
However, we also found that the particular effect of environment is strongly dependent on the species.
We proposed two altenative hypotheses of how environmental stress may affect degree and we found evidence for both of them.
Species with a large number of partners in low-stress communities were more likely to have a negative relationship and hence reduce the number of partners as stress increases. 
Contrastingly, species in our datasets with a small number of partners in low-stress communities were more likely to have a larger number of partners in more stressful communities.

Our results suggest that changes in community composition are indeed the primmary channel through which environment determines changes interaction probability. 
However, they also show that, for a large number of species, the environment may also play a substantial role on determining their realised (Eltonian) niche. 
While previous research has recognised that environmental factors may help explain the changes in network structure along environmental gradients that cannot be explained by community composition [@tylianakis_habitat_2007], how these two factors were linked had been elusive so far [@gravel_bringing_2018]. 
We believe that part of this difficulty could have arisen because species, and ultimately network structre, can respond multiple, and contrasting, ways depending on the particular bioclimatic variable being examined (e.g. temperature or precipitation).
Using stress to summarise the effect on species of multiple environmental gradients allowed us to detect a clear signal of the environment in species' interaction patterns.

Although both niche and neutral processes are relevant at determining species interactions, our model suggests that niche processes may be the predominant mechanism through which the environment *systematically* affect species degree. 
First, it is unlikely that environmental suitability is correlated to local species abundances [@pearce_practical_2001; @sagarin_moving_2006]. 
Second, even if there is a relationship between suitability and abundances, a particular environmental gradient could have a positive effect in the abundance some species and a negative effect in others. 
Indeed, we show that within a community there is a wide range of suitability values, even for the relatively limited number of species we were able to include in our analyis.

Recent research suggest that species are continuoisly chaning there interaction partners wherever environmental conditions change in space of time [@raimundo_adaptive_2018].
So far it appears that this rewiring is primarly driven by generalist species [@ponisio_opportunistic_2017; @burkle_plant-pollinator_2013], presumably because generalist species are less sensitive to trait matching of their interaction partners [@caradonna_interaction_2017].
Our results add two important nuances to these findings. 
First, because generalists seem to focus on a smaller number of partners as environmental conditions deteriorate, we show that trait matching might still play a role in determining the interactions of generalist species. 
Second, we demonstrate that rewiring is not exclusive to generalists.
At least a fraction of the species that appear to be specialist in their communities might be as flexible, if not more, than generalist species, effectively behaving as facultative generalists in the face of environmental change. 
These flexible specialists might, therefore, have a more significant role in network persistence than previously expected.

While our model detected how the environment affects interaction of two types of species, generalists and flexible specialists, there is a third group that remained invisible for our model but has important implications for network persistence and stability.
It has been shown that species that are able to flexibly vary their interaction partners and their role in the network are more likely to persist in their community as environmental conditions vary [@gaiarsa_interaction_2019].
We propose this third group of is composed by specialists that are constrained to interact with parterns of high trait-matching and therefore were not likely to be found in more than one community. 
If species that are not flexible are unlikely to persist over temporal or spatial environmental gradients we can expect specialised communities that are highly constrained by trait-matching [like some plant-hummingbird networks; @vizentin-bugoni_processes_2014; @maruyama_morphological_2014] to be far more vulnerable to increased climate change-induced environmental stress and habitat degradation than communities where role and interaction flexibility are more prevalent.

Similarly, if the patterns we see in our models have also played a role during the evolutionary history of pollination communities, our results also help explain why only a small fraction of plant-pollinator interactions show a strong signature of deep co-evolutionary history.
The increasses on stress that species are predicted to experience due to rapid environmental change might further erode the co-evolutionary history of specialist species.
Communities as a whole might be set in a trayectory of even more diffuse co-evolution.
For specialist species, at least, the longer term benefits of being able to interact with multiple partners might be more important than the shorter term benefits of interacting with partners of high trait matching.

>>>>

Species interactions are determined in part by niche processes (the matching of traits) and partly by neutral processes (more abundant species are more likely to encounter each other and, thus, interact). 
The environment can influence both of these processes. 
It is, therefore, not surprising that, despite limitations on the spatial extent or the number of environmental gradients considered, multiple studies have been able to show how changes to interactions can be related to environmental change [@tylianakis_ecological_2017].
For instance, some studies suggest that the strength of some trophic interactions, like predation [@mckinnon_lower_2010; @vucic-pestic_warming_2011] and herbivory [@baskett_latitudinal_2018], can increase with temperature but might decrease with precipitation [@pires_predicted_2016]. 
Some other studies, however, have shown either no effect (on average) or non-linear effects of temperature or precipitation on plant-pollinator interactions [@devoto_patterns_2005; @gravel_bringing_2018].
Overall, while it looks clear that pairwise interactions respond to environmental drivers, there is high variability in the response [@tylianakis_global_2008].

@@

Here, we investigate whether, and how, environmental stress can systematically affect species degree. 
Our main aim is to test the two competing hypotheses that relate environmental stress and species degree and investigate whether this changes across species or between trophic guilds. 
We propose that specialist species can become "facultative" generalists to reduce their vulnerability to the absence of preferred partners [for example, when variations in climate decouple phenologies; @benadi_specialization_2014]. 
We therefore also expect that as environmental stress increases species with a relatively small number of partners are more likely to engage with more partners and broaden their trophic niche. 
Species with a large number of partners, on the other hand, should have a larger pool of available partners and might, therefore, be more likely to narrow their niche under environmental stress by focusing on the most beneficial partners. 
Importantly, when testing these hypotheses, we control for the potential effects of the environment in community composition and the size of the species fundamental niche, both from an Eltonian (interactions) and Grinellian (environment) perspective.

@@

We retrieved plant-pollinator networks from the Web of Life database [@fortuna_web_2014]. 
This database contains datasets originating from `r length(interaction_citations$BIBTEXKEY)` studies published in the primary literature between `r int_citations_year_range[1]` and `r int_citations_year_range[2]`. 
Calculating the environmental stress of species in their community and their Eltonian niche breadth required us to reduce both the taxonomic and distributional/locational uncertainty.
A critical step towards reducing this uncertainty is to ensure that the names used to identify species are valid and unambiguous, which in turn allow us to obtain further information from biological databases and accurately match species across studies. 
Therefore, our first step was to ensure consistent spelling and standardisation of species names synonyms (See Supplementary Methods).
This ensured that the matching of species across studies was as accurate as possible. 
The cleaning process resulted on a total of `r format(stats_clean_int$pla_id, big.mark = ",")` plants and `r format(stats_clean_int$ani_id, big.mark = ",")` pollinator species distributed across `r format(stats_clean_int$loc_id)` locations arround the globe (Figure \@ref(fig:fig-worldmap), \@ref(fig:fig-distribution-species-mult-locations)).

@@

We retrieved `r round(sum(n_occurrences$N)/1000000, digits = 1)` million occurrences from the Global Biodiversity Information Facility (GBIF; https://www.gbif.org). 
Issues with data quality are a central issue hampering the use of publicly available species occurrence GBIF data in ecology and biogeography [@jetz_essential_2019]. 
We, therefore we followed a series of filters and geographic heuristics to correct or remove erroneous and imprecise referencing records [See supplementary methods; @zizka_coordinatecleaner_2019] which allowed us to identify and remove `r round(sum(n_dirty_occurrences$N)/1000000, digits = 1)` million problematic occurrences from further analysis.
We integrated the occurrences from our plant-pollinator communities to the cleaned occurrences retrieved from GBIF.

@@

Sets of occurrence data tend to be spatially aggregated due to sample bias (tendency to collect close to cities, certain countries). 
Spatial autocorrelation arises in ecological data because geographically clumped records tend to be more similar, in physical characteristics and/or species abundances, than are pairs of locations that are farther apart. 
To account for such spatial dependency in occurrence data, if a species had more than one occurrence records within one of the cells of the bioclimatic raster, we only included one of the occurrence records. 
We did this to avoid giving more weight to areas with a high number of occurrences, a common scenario in occurrence records collected opportunistically as the ones we use here. 
In this step we removed `r scales::percent(percent_thinned)` of the occurrences which resulted in a total of `r round(sum(n_thinned_occurrences$N)/1000000, digits = 1)` million occurrences used in our niche analysis.

@@

We then used a set of bayesian multilevel models to evaluate the impact of environmental suitability on the number of partners a species has. 
Specifically, we use the normalised degree of species as our response variable; this is, the number of species it interacts with given the number of species in the opposite guild [@martin_gonzalez_centrality_2010]. 
The normalised degree was modelled using a logit link function, and a binomial distribution in which the number of species interacts with is the number of successes, and the number of species in the opposite guild is the number of trials. 
We are aware that whether species interact or not is not a Bernoulli process as species interactions are not strictly independent from each other. 
However, a binomial distribution allows us to account for the differences in species richness across communities indirectly. 
Importantly, however, results are qualitatively similar when we model species degree directly using a Poisson distribution and a logarithmic link function.

We evaluate four models to assess the relative importance of suitability. 
A first model, our baseline model, included three population-level predictors and two grouping levels, species and the community.
The population-level predictors in the baseline model, commonly called fixed effects, were the habitat suitability, the species guild (plant or a pollinator), and its number of known possible partners.
We included the number of known possible partners as a predictor in our models as it allows us to control for the environmental effects on species co-occurrence. 
We calculate this metric by determining the number of partners with which the species is known to interact in any other community.
Controlling for the number of potential partners makes our model a particularly stringent test of our environmental stress hypotheses because this variable could explain a large proportion of variance. 
Often the potential and the actual number of partners is the same or very close to each other, especially for rare species present only in a few communities.
As we were interested in understanding whether the effect of habitat suitability is conditional on the species guild (plant or pollinator), we, therefore, included guild and its interaction with suitability in the model. 
We allowed the intercept of degree and slope of the suitability-degree relationship to vary among species. 
This allowed us to investigate two questions. 
First, it allows us to inspect the extent to which suitability is a population or a group level effect. 
Second, by investigating the correlation between the intercept and the slope as a model parameter, it allowed us to inspect the extent by which species with a small or large number of interactions respond to increasing levels of environmental stress. 
To account for unmeasured differences between communities, like sampling effort, sampling method, or diversity, we also calculated an intercept for each community in our study.
To facilitate model interpretation and convergence, we scaled all continuous variables to have a mean of zero and a unit standard deviation.

@@

After performing our sensitivity analysis, we found that we need roughly `r readd(min_occurrences_factor)` independent occurrences for each community for which we calculated a suitability value in order to obtain a mean absolute error below `r config$min_suitability_error` (Fig. \@ref(fig:fig-sensitivity-analysis)). 
We, therefore, removed from further analyses `r n_spp_not_enough_occ` species for which we did not have enough occurrences to obtain robust estimates.
When inspecting the suitability values of the analysed species, we found that most communities included species for which habitat suitability was low and species for which it was high (Fig. \@ref(fig:fig-median-suitability)).

@@

Our models performed relatively well.
The bayesian R-squared for our baseline model was `r round(bayesian_r2_baseline[1], digits = 2)`, which indicates our models were able to capture a large proportion of the variability on the data. 
Overall we found that environmental suitability does not show a consistent pattern across species. 
Indeed, when looking at the population level effects, suitability has virtually no relationship with the normalised degree, neither for plants or pollinators (Figure \@ref(fig:fig-conditional-effects)b). 
However, suitability is an important predictor as the WAIC difference between our baseline model and that that did not include suitability was `r round(waic_diff_suitability)` ± `r round(waic_se_diff_suitability)` (Table \@ref(tab:tab-model-comparison)).
This apparent discrepancy can be explained by the variability of the suitability-degree relationship across species.

@@

fig_conditional_effects_legend <- 
  paste0("Conditional effects of predictors in our baseline model. The predicted values of the number of interacting species are based on a hypothetical community with ",
         median_n_pla, " plants and ", 
         median_n_ani, " pollinators",
         ". These values correspond to the median number of species in each guild, respectively. In each panel, we condition on the mean values of all other predictors in the model, mean values for each predictor are indicated with a vertical dashed line. For model fitting, we scaled all predictors to have a mean of zero and unit variance. However, with the exception of environmental niche size, here we show the unscaled predictors to facilitate interpretation. To illustrate the uncertainty around the fitted estimates, we plot the fits of 100 independent draws from the posterior distribution. The thick lines indicate the mean values of the response distribution. As there was no interaction between guild and generality or the number of possible interactions, for these two predictors, we only show conditional the conditional effect of pollinators.")

@@

While for some species there is a strong negative relationship between suitability and normalised degree, for some others with a strong positive relationship (Figure \@ref(fig:fig-random-effects)a). 
Interestingly, the slope of this relationship was negatively correlated with the species' intercept in the model (Figure \@ref(fig:fig-random-effects)b). 
The mean correlation coefficient was `r round(random_cor$Estimate, 2)` [`r round(random_cor$Q2.5, 2)`, `r round(random_cor$Q97.5, 2)`].
In other words, the slope of the suitability-degree relationship was more likely to be positive for species with a smaller number of partners in lower-suitability and negative for species with a larger number of partners in lower-suitability conditions.

@@

fig_random_effects_legend <- 
  paste0("Species-level effects of suitability. (a) Conditional effect of suitability for individual species. To facilitate visualization we show only species for which there is suitability information in at least six communities (", 
         n_random_pla, " plants and ", 
         n_random_ani, " pollinators). As in the previous figure, fitted values assume a hypothetical community of median size. In each panel, we highlight two species for which the relationship between environmental suitability and the normalised degree was particularly strong. ", 
         "(b) The correlation between the species' intercept and the species' slope of suitability was negatively correlated.")

@@

As expected, we found a strong and positive relationship between the number of possible interactions and the number of realised interactions in the community. 
There was a large difference on WAIC between the model that included this predictor and that that excluded it. 
This result indicates that the availability of potential partners accounts for a large proportion of the variability on species degree.  
Importantly, our findings of the suitability-degree relationship were qualitatively unchanged whether this variable was included or not.

The group-level variation among communities was larger than that among species which further indicates the importance of the local context when determining species degree.
Specifically, the standard deviation (in the parameters scale) of the community intercepts was `r round(sd_loc_id$Estimate, 2)` [`r round(sd_loc_id$Q2.5, 2)`, `r round(sd_loc_id$Q97.5, 2)`] while the standard deviation of the species intercept was `r round(sd_org_id$Estimate, 2)` [`r round(sd_org_id$Q2.5, 2)`, `r round(sd_org_id$Q97.5, 2)`], and that of the species' suitability slope was `r round(sd_org_id_suitability$Estimate, 2)` [`r round(sd_org_id_suitability$Q2.5, 2)`, `r round(sd_org_id_suitability$Q97.5, 2)`] (95% credible intervals shown within square brakets).

@@

We set out to explore whether and how environmental stress can systematically affect species degree. 
After accounting for the pool of potential partners, we found that indeed environmental conditions contribute to determining whether a species is a generalist or a specialist in their community. 
However, we also found that the particular effect of the environment is strongly dependent on the species.
We proposed two alternative hypotheses of how environmental stress may affect the degree, and we found evidence for both of them.
Species with a large number of partners in low-stress communities were more likely to have a negative relationship and hence reduce the number of partners as stress increases. 
Contrastingly, species in our datasets with a small number of partners in low-stress communities were more likely to have a larger number of partners in more stressful communities.
In summary, environmental stress pushes species that are flexible enough to change their interaction partners towards intermediate levels of degree, a so-called "regression towards the mean".

Our results suggest that changes in community composition are indeed the primary channel through which the environment determines changes interaction probability. 
However, they also show that, for a large number of species, the environment may also play a substantial role in determining their realised (Eltonian) niche. 
While previous research has recognised that environmental factors may help explain the changes in network structure along environmental gradients that cannot be explained by community composition [@tylianakis_habitat_2007], how these two factors were linked had been elusive so far [@gravel_bringing_2018]. 
We believe that part of this difficulty could have arisen because species, and ultimately network structure, can respond in multiple and contrasting, ways depending on the particular bioclimatic variable being examined (e.g. temperature or precipitation).
Using stress to summarise the effect on species of multiple environmental gradients allowed us to detect a clear signal of the environment in species' interaction patterns.

Although both niche and neutral processes are relevant at determining species interactions, our model suggests that niche processes may be the predominant mechanism through which the environment *systematically* affect species degree. 
First, it is unlikely that environmental suitability is correlated to local species abundances [@pearce_practical_2001; @sagarin_moving_2006]. 
Second, even if there is a relationship between suitability and abundances, a particular environmental gradient could have a positive effect on the abundance of some species and a negative effect on others. 
Indeed, we show that within a community there is a wide range of suitability values, even for the relatively limited number of species we were able to include in our analysis.

Recent research suggests that species are continuously changing their interaction partners wherever environmental conditions change in space or time [@raimundo_adaptive_2018].
So far it appears that this rewiring is primarily driven by generalist species [@ponisio_opportunistic_2017; @burkle_plant-pollinator_2013], presumably because generalist species are less sensitive to trait matching of their interaction partners [@caradonna_interaction_2017].
Our results add two important nuances to these findings. 
First, because generalists seem to focus on a smaller number of partners as environmental conditions deteriorate, we show that trait matching might still play a role in determining the interactions of generalist species. 
Second, we demonstrate that rewiring is not exclusive to generalists.
At least a fraction of the species that appear to be specialist in their communities might be as flexible, if not more, than generalist species, effectively behaving as facultative generalists in the face of environmental change. 
These flexible specialists might, therefore, have a more significant role in network persistence than previously expected.

While our model detected how the environment affects the interactions of two types of species, generalists and flexible specialists, there is a third group that remained invisible for our model but has important implications for network persistence and stability.
It has been shown that species that are able to vary their interaction partners flexibly and their role in the network are more likely to persist in their community as environmental conditions vary [@gaiarsa_interaction_2019].
We propose this third group of is composed of specialists that are constrained to interact with partners of high trait-matching and therefore were not likely to be found in more than one community. 
If species that are not flexible are unlikely to persist over temporal or spatial environmental gradients we can expect specialised communities that are highly constrained by trait-matching [like some plant-hummingbird networks; @vizentin-bugoni_processes_2014; @maruyama_morphological_2014] to be far more vulnerable to increased climate change-induced environmental stress and habitat degradation than communities where role and interaction flexibility are more prevalent.

Similarly, if the patterns we see in our models have also played a role during the evolutionary history of pollination communities, our results also help explain why only a small fraction of plant-pollinator interactions show a strong signature of deep co-evolutionary history [@hutchinson_cophylogenetic_2017].
The increases on the stress that species are predicted to experience due to rapid environmental change might further erode the co-evolutionary history of specialist species.
Communities as a whole might be set in a trajectory of even more diffuse co-evolution.
For specialist species, at least, the longer-term benefits of being able to interact with multiple partners might be more important than the shorter-term benefits of interacting with partners of high trait matching.
